On the error performance of the $A_n$ lattices

### 2011-11-16 paper.tex
Added minor comments from Emanuele.
Fixed a couple of minor things: (1) Logical ordering of description in first paragraph of Section II. (2) argmax display in Eqn.(1)

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis. Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix. The set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.

@@

%For each lattice point $\ybf \in \Lambda$ there exists a Voronoi cell containing those point from $\reals^n$ nearer to $\ybf$ that any other lattice point.  Due to the structure of the lattice the Voronoi cell of each lattice point is the same shape.  Denote by $\vor(\Lambda)$ the Vornoi cell of the lattice point at the origin. NEAREST POINT PROBLEM.
\newcommand{\calX}{\mathcal X}
Lattices can be used to construct \emph{signal space codes}.  A signal space code $\calX$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.  Each vector in $\calX$ is called a \emph{codeword} and represents a particular signal.  The number of codewords is denoted by $\abs{\calX}$.  If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calX}$ bits per codeword. The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the AWGN channel the received signal takes the form
\[
\ybf = \xbf + \wbf
\]
where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.  If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
\begin{equation}\label{eq:mldecoder}
\hat{\xbf} = \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
\end{equation}
That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.  Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
\[
P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ),
\]
and the probability of error is
\[
P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
\hat{\xbf} = \xbf ).
\]
%The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.  If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.  Another way of stating this is that we can choose close to $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.

>>>>

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional sub-space containing the lattice $\Lambda$.

@@

%For each lattice point $\ybf \in \Lambda$ there exists a Voronoi cell containing those point from $\reals^n$ nearer to $\ybf$ that any other lattice point.  Due to the structure of the lattice the Voronoi cell of each lattice point is the same shape.  Denote by $\vor(\Lambda)$ the Vornoi cell of the lattice point at the origin. NEAREST POINT PROBLEM.
\newcommand{\calX}{\mathcal X}
Lattices can be used to construct \emph{signal space codes}.  A signal space code $\calX$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.  Each vector in $\calX$ is called a \emph{codeword} and represents a particular signal.  The number of codewords is denoted by $\abs{\calX}$.  If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calX}$ bits per codeword. The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the AWGN channel the received signal takes the form
\[
\ybf = \xbf + \wbf
\]
where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.  If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
\begin{equation}\label{eq:mldecoder}
\hat{\xbf} = \underset{\xbf \in \calX}{\operatorname{argmin}}\;\| \ybf - \xbf \|^2.
%\hat{\xbf} &=& \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
\end{equation}
That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.  Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
\[
P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ),
\]
and the probability of error is
\[
P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
\hat{\xbf} = \xbf ).
\]
%The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.  If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.  Another way of stating this is that we can choose close to $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.
